The Castro AMR Simulation Code: Current and Future Developments

### 2019-10-02 astronum_2019_Castro/astronum-castro-2019.tex

The SDC algorithm used in \castro\ follows the ideas of
\cite{dutt:2000,minion:2003}, and uses low order explicit advection
and implicit reaction updates in a correction equation that when
applied iteratively achieves high-order time-accuracy.  In an operator
split implementation, reaction and burning proceed without knowing
about the other process.  In contrast, the update in the SDC
formulation explicitly couples the processes together.  In \castro\ we
consider reactions and hydrodynamics, and write our conservative
system as:
\begin{equation}
\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
\end{equation}
where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
advective term (the divergence of the hydrodynamic fluxes along with
hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
The update is done at specific temporal nodes, the number of locations
of which are picked to give the desired temporal accuracy.  The update
from one time node $m$ to $m+1$ appears as:
    \begin{align}
      \label{eq:sdc:general}
      \avg{\Uc}^{{m+1},(k+1)} = \avg{\Uc}^{m,(k+1)}
            &+ \delta t_m \left [ \avg{\Advs{\Uc}}^{m,(k+1)} - \avg{\Advs{\Uc}}^{m,(k)} \right ] \nonumber \\
            &+ \delta t_m \left [ \avg{\Rbs{\Uc}}^{{m+1},(k+1)} - \avg{\Rbs{\Uc}}^{{m+1},(k)} \right ] \nonumber \\
            &+ \int_{t^m}^{t^{m+1}} dt  \left (\avg{\Advs{\Uc}}^{(k)} + \avg{\Rbs{\Uc}}^{(k)}\right)
   \end{align}
We use a finite-volume formalism, so $\avg{\Uc}$ is the spatial average of the
conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ \MarginPar{I think this should be $\avg{\Rbs{\Uc}}^{m+1}$?}is
the average of the reactive source at time $t^{m+1}$.  The second
superscript on each term, $(k)$ or $(k+1)$, is the iteration counter.  This
update is an implicit equation for the new state,
$\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is an integral
over the sources constructed from the previous iteration's values at
each time node.  Each iteration of the SDC method increases the
temporal order of accuracy by one, up to the order of accuracy with
which the integral is constructed.

>>>>

The SDC algorithm used in \castro\ follows the ideas of
\cite{dutt:2000,minion:2003}, and uses low order explicit advection
and implicit reaction updates in a correction equation that when
applied iteratively achieves high-order time-accuracy.  In an operator
split implementation, reaction and burning proceed without knowing
about the other process.  In contrast, the update in the SDC
formulation explicitly couples the processes together.  In \castro\ we
consider reactions and hydrodynamics, and write our conservative
system as:
\begin{equation}
\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
\end{equation}
where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
advective term (the divergence of the hydrodynamic fluxes along with
hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
The update is done at specific temporal nodes, the number of locations
of which are picked to give the desired temporal accuracy.  The update
from one time node $m$ to $m+1$ appears as:
    \begin{align}
      \label{eq:sdc:general}
      \avg{\Uc}^{{m+1},(k+1)} = \avg{\Uc}^{m,(k+1)}
            &+ \delta t_m \left [ \avg{\Advs{\Uc}}^{m,(k+1)} - \avg{\Advs{\Uc}}^{m,(k)} \right ] \nonumber \\
            &+ \delta t_m \left [ \avg{\Rbs{\Uc}}^{{m+1},(k+1)} - \avg{\Rbs{\Uc}}^{{m+1},(k)} \right ] \nonumber \\
            &+ \int_{t^m}^{t^{m+1}} dt  \left (\avg{\Advs{\Uc}}^{(k)} + \avg{\Rbs{\Uc}}^{(k)}\right)
   \end{align}
We use a finite-volume formalism, so $\avg{\Uc}$ is the spatial average of the
conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{m+1}$
the average of the reactive source at time $t^{m+1}$.  The second
superscript on each term, $(k)$ or $(k+1)$, is the iteration counter.  This
update is an implicit equation for the new state,
$\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is an integral
over the sources constructed from the previous iteration's values at
each time node.  Each iteration of the SDC method increases the
temporal order of accuracy by one, up to the order of accuracy with
which the integral is constructed.
